---
title: "GenderizeR"
bibliography: references.bib
link-citations: yes
---


  
```{r, globalsettings, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"), cache.lazy = FALSE)
options(width = 100) 
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE, message=FALSE}
# install.packages("remotes")
#remotes::install_github("rlesur/klippy")
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


----

# Goal

This file is used to determine gender based on first name for our sample of sociologists and political scientists. 



# Custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R ([source](https://vbaliga.github.io/verify-that-r-packages-are-installed-and-loaded/)).  
- `fsave`: Save to processed data in repository  
- `fload`: To load the files back after an `fsave`  
- `fshowdf`: To print objects (tibbles / data.frame) nicely on screen in .rmd  


```{r customfunctions, results='hide'}

rm(list = ls())

fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}

fsave <- function(x, file=NULL, location="./data/processed/") {
  ifelse(!dir.exists("data"), dir.create("data"), FALSE)
  ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
  if (is.null(file)) file= deparse(substitute(x))
  datename <- substr(gsub("[:-]", "", Sys.time()), 1,8)  
  totalname <- paste(location, datename, file, ".rda", sep="")
  save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
  load(filename)
  get(ls()[ls() != "filename"])
}

fshowdf <-  function(x, ...) {
  knitr::kable(x, digits=2, "html", ...) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::scroll_box(width="100%", height= "300px")
} 

```


---  

# Packages

- `genderizeR`: genderize package adapted to use in R, maps first names to gender


```{r, results='hide', message=FALSE}

packages = c("genderizeR", "tidyverse")

fpackage.check(packages)

```


--- 

# Input

This is the file you made in the previous step. 

If something went wrong, you can download the file here: 

`r xfun::embed_file("./data/processed/20230627df_names.rda")`  

Put it in the correct folder: './data/processed' 

```{r}

df <- fload(file="./data/processed/20230627df_names.rda")

```



---



# Preparing the dataframe

We only need to check each unique first name once, so therefore we extract all unique names first. 
We use the `distinct` function from tidyverse to recognize and then `select` only unique entries in the first name object

```{r}

firstname <- df %>% 
  distinct(firstname) %>% #only unique
  select(firstname) %>% #only this variable
  filter(firstname != "") %>% #remove empty
  pull(firstname) #save as chr


```



# GenderizeR 

It is possible to use specific databases of GenderizeR by country. In that case, you can add the 'country' argument in genderizeAPI. We will now use the worldwide database, which this function defaults to when you don't add a country argument. 

As a besides, in our 'real' project, we scraped the gender for Dutch names from this [website](https://nvb.meertens.knaw.nl/). We only used genderizeR as check. 

Let us scrape one name first  

```{r}
test <- genderizeAPI(firstname[1])
test$response
```
Okay, now just loop. ;-)

First we scrape...but watch out, you will probably run into rate limits. 


```{r, eval=FALSE}
length(firstname)
df_genderizer <- data.frame(matrix(ncol = 4, nrow = length(firstname)))
colnames(df_genderizer) <- c('count', 'name', 'gender', 'probability')
for (i in 1:length(firstname)) {
  Sys.sleep(0.1)
  df_genderizer[i,] <- genderizeAPI(firstname[i])$response
  }
colnames(df_genderizer) <- c('count', 'name', 'gender', 'probability')
df_genderizer$name <- firstname #so we know which names are not found. 
```



<!--- 
Intermittent save (for us, remove before WS)
```{r, eval=FALSE}
save(genderizer, file = "data/processed/genderizer_raw.rda")

save(df_genderizer, file = "data/processed/df_genderizer.rda")

```
---> 

You may get stuck along the way. We therefore the objects for you. 

You can download the files here: 

`r xfun::embed_file("./data/processed/genderizer_raw.rda")`  

`r xfun::embed_file("./data/processed/df_genderizer.rda")`  

Put it in the correct folder: './data/processed' 

```{r, echo=FALSE}
load("./data/processed/genderizer_raw.rda")

load("./data/processed/df_genderizer.rda")
```


Decision moment! 

We retain only names where we can be pretty sure about the gender. If the name is around 50/50 male- or female-typed, we cannot assign the gender with confidence.

Our cut-off point is at .90, meaning we can be 90% confident that a name is either male or female. 

To be honest, I think it would be a better strategy not to dummify but simply use the probability score as variable in your final model. 

```{r}
df_genderizer <- df_genderizer[df_genderizer$probability>.9, ]


df_genderizer <- subset(df_genderizer, select=c("name", "gender"))

fshowdf(df_genderizer)
```



# Merging the genderizer info to the dataframe with names

```{r}

df %>%
  left_join(df_genderizer, by=c("firstname"="name")) -> df_gender

fshowdf(df_gender)
```


# Saving the dataframe with gender

```{r, eval=FALSE}

fsave(df_gender)

```

