---
title: "GenderizeR"
bibliography: references.bib
link-citations: yes
---


  
```{r, globalsettings, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"), cache.lazy = FALSE)
options(width = 100) 
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE, message=FALSE}
# install.packages("remotes")
#remotes::install_github("rlesur/klippy")
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


----

# Goal

This file is used to determine gender based on first name for our sample of sociologists and political scientists. 



# Custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R ([source](https://vbaliga.github.io/verify-that-r-packages-are-installed-and-loaded/)).  
- `fsave`: Save to processed data in repository  
- `fload`: To load the files back after an `fsave`  
- `fshowdf`: To print objects (tibbles / data.frame) nicely on screen in .rmd  


```{r customfunctions, results='hide'}

rm(list = ls())

fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}

fsave <- function(x, file=NULL, location="./data/processed/") {
  ifelse(!dir.exists("data"), dir.create("data"), FALSE)
  ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
  if (is.null(file)) file= deparse(substitute(x))
  datename <- substr(gsub("[:-]", "", Sys.time()), 1,8)  
  totalname <- paste(location, datename, file, ".rda", sep="")
  save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
  load(filename)
  get(ls()[ls() != "filename"])
}

fshowdf <-  function(x, ...) {
  knitr::kable(x, digits=2, "html", ...) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::scroll_box(width="100%", height= "300px")
} 

```


---  

# Packages

- `genderizeR`: genderize package adapted to use in R, maps first names to gender


```{r, results='hide', message=FALSE}

packages = c("genderizeR", "tidyverse")

fpackage.check(packages)

```


--- 

# Input

This is the file you made in the previous step. 

If something went wrong, you can download the file here: 

`r xfun::embed_file("./data/processed/20230627df_names.rda")`  

Put it in the correct folder: './data/processed' 

```{r}

df <- fload(file="./data/processed/20230627df_names.rda")

```



---



# Preparing the dataframe

We only need to check each unique first name once, so therefore we extract all unique names first. 

```{r}
# We use the distinct function from tidyverse to recognize and then select only unique entries in the first name object
uniquenames <- df %>% distinct(firstname) %>% select(firstname)
uniquenames <- uniquenames[uniquenames$firstname!="", ] # transform to character vector


# create empty lists to store the genderizeR info
genderizer <- list()

names <- list()
count <- list()
gender <- list()
prob <- list()
```



# GenderizeR 

It is possible to use specific databases of GenderizeR by country. In that case, you can add the 'country' argument in genderizeAPI. We will now use the worldwide database, which this function defaults to when you don't add a country argument. 

Let us scrape one name first  

```{r}
genderizeAPI(uniquenames[1])
```
Okay, now just loop. ;-)

First we scrape...


```{r, eval=FALSE}
for (i in 1:length(uniquenames)) {
  
  Sys.sleep(0.1)
  
  genderizer[[i]] <- genderizeAPI(uniquenames[i])[[1]]
  
}
```

... then we fill in names for which no info is found: count to 0, gender and probability to NA, but retain the name for reference. 

```{r, eval=FALSE}
for (i in 1:length(genderizer)) {
  
  if (nrow(genderizer[[i]])==0) {
    
    genderizer[[i]][1, c("count")] <- 0
    genderizer[[i]][1, c("gender")] <- NA
    genderizer[[i]][1, c("name")] <- uniquenames[i]
    genderizer[[i]][1, c("probability")] <- NA
    
  }
}
```

And we convert from list to dataframe structure:  

```{r, eval=FALSE}

for (i in 1:length(genderizer)) {
  
  names[[i]] <- data.frame(genderizer[[i]])$name
  count[[i]] <- data.frame(genderizer[[i]])$count
  gender[[i]] <- data.frame(genderizer[[i]])$gender
  prob[[i]] <- data.frame(genderizer[[i]])$probability

}


df_genderizer <- do.call(rbind, Map(data.frame, name = names, count = count, gender = gender, probability = prob)) 


# setting count = 0 for names that cannot be found in the for-loop does not work well
df_genderizer$count <- ifelse(is.na(df_genderizer$count), 0, df_genderizer$count)
```


<!--- 
Intermittent save (for us, remove before WS)
```{r, eval=FALSE}
save(genderizer, file = "data/processed/genderizer_raw.rda")

save(df_genderizer, file = "data/processed/df_genderizer.rda")

```
---> 

You may get stuck along the way. We therefore the objects for you. 

You can download the files here: 

`r xfun::embed_file("./data/processed/genderizer_raw.rda")`  

`r xfun::embed_file("./data/processed/df_genderizer.rda")`  

Put it in the correct folder: './data/processed' 

```{r, echo=FALSE}
load("./data/processed/genderizer_raw.rda")

load("./data/processed/df_genderizer.rda")
```


Decision moment! 

We retain only names where we can be pretty sure about the gender. If the name is around 50/50 male- or female-typed, we cannot assign the gender with confidence.

Our cut-off point is at .90, meaning we can be 90% confident that a name is either male or female.

```{r}
df_genderizer <- df_genderizer[df_genderizer$probability>.9, ]


df_genderizer <- subset(df_genderizer, select=c("name", "gender"))

fshowdf(df_genderizer)
```



# Merging the genderizer info to the dataframe with names

```{r}

df %>%
  left_join(df_genderizer, by=c("firstname"="name")) -> df_gender

fshowdf(df_gender)
```


# Saving the dataframe with gender

```{r, eval=FALSE}

fsave(df_gender)

```

