---
title: "Sunbelt 2024 Workshop: Webscraping"
author: 
  - '**AUTHORS:**'
  - '[Jochem Tolsma](https://www.jochemtolsma.nl) - Radboud University / University of Groningen, the Netherlands'
  - '[Rob Franken](https://robfranken.net/) - Utrecht University, the Netherlands'
  - '[Dan Cowen](https://www.rug.nl/staff/d.r.cowen/) - University of Groningen, the Netherlands'
  - '[Anne Maaike Mulders](https://www.ru.nl/en/people/mulders-a) - Radboud University, the Netherlands'
  - '[Bas Hofstra](https://www.bashofstra.com/) - Radboud University, the Netherlands'
date: "Last compiled on `r format(Sys.Date(), format='%d %B %Y')`"
bibliography: references.bib
link-citations: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r, globalsettings, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"), cache.lazy = FALSE)
options(width = 100) 
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE, message=FALSE}
# install.packages("remotes")
#remotes::install_github("rlesur/klippy")
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

------------------------------------------------------------------------

# Introduction

```{=html}
<script>
function myFunction() {

            var btn = document.getElementById("myButton");
            //to make it fancier
            if (btn.value == "Click to hide code") {
                btn.value = "Read more";
                btn.innerHTML = "Read more";
            }
            else {
                btn.value = "Click to hide code";
                btn.innerHTML = "Read less";
            }
            //this is what you're looking for
            var x = document.getElementById("myDIV");
            if (x.style.display === "none") {
                x.style.display = "block";
            } else {
                x.style.display = "none";
            }
        }
          
</script>
```
Welcome to the website of the Sunbelt 2024 Workshop: **Webscraping
Scientific Co-publishing Networks**

<br>

In this workshop you will collect and then describe academic
departmental co-publishing networks. You will learn to webscrape
scientific metadata of scientific university and departmental websites
(via R packages like rvest and RSelenium), assign name-based gender and
ethnicity signals, retrieve scholars' publications (via e.g. Google
Scholar and OpenAlex), and construct (longitudinal) co-publishing
networks.

In our description of these networks, we will focus on the degree of
segregation. You will answer questions such as: Are co-publishing
networks segregated by scientific success and/or gender? And is
segregation in co-publishing networks related to university/departmental
characteristics?

For each step, we provide clear (proof-of-principle) coding examples and
output data, ensuring you will not get stuck along the way. Depending on
your skills and progress, you might want to collect and describe your
own chosen universities or departments.

This workshop can be followed as a standalone workshop but in our second
workshop 'Analyzing the Structure and Evolution of Scientific
Co-publishing Networks' we will describe and analyze the same type of
webscraped co-publishing network data in more detail by employing
RSiena.

You will keep track of your work via a labjournal on GitHub. We will
help you setting up this labjournal in the week(s) before the workshop.

Prerequisites:

-   Intermediate familiarity with working in R and using R Markdown\
-   A beginner's understanding of SNA\
-   Entry-level of git, and GitHub

```{=html}
<!--

<button class="button1" onclick="myFunction()" id="myButton" value="Click To Open Instructions">

Read more

</button>

::: {#myDIV style="display:none;"}
<br>

more text here

<br>
:::

--->
```

------------------------------------------------------------------------

# Getting started

## LabJournal

During the course, students will journal their work and assignments in
their custom lab journal. A template lab journal can be found on
[GitHub](https://github.com/robfranken/LabJournal). Here, you find how
to get started.

<!-- ## Discord -->

<!-- Please join the [discord channel](https://discord.gg/ACcYU8nu). Use the -->

<!-- channel to chat, ask questions and share (small) files. -->

------------------------------------------------------------------------

# Program

The program of this workshop will be as follows:

`r colorize("Introduction", "red")`\
*Time: 9:00am - 9:30am*

-   **all**
    -   \- round of introductions
-   **Jochem, Rob, Dan**
    -   Our goals for today
    -   Research Questions based on Scientific Co-publishing Networks
    -   Data requirements for RQs\

`r colorize("Lab Journal", "red")`\
*Time: 9:30am - 10:30am*

-   **Rob**
    -   \- [Lab Journal](https://github.com/robfranken/LabJournal)
    -   Some Git pointers
    -    Some RMarkdown pointers

`r colorize("BREAK", "red")`\
*Time: 10:30am - 10:45am*

`r colorize("Webscraping-scholars Part 1", "red")`\
*Time: 10:45am - 12:30pm*

-   **Jochem**
    -   Webscraping - [Lecture](./presentations/insna_science_data.pdf)
-   **Dan**
    -   Webscraping - [Tutorial](RUG_scholar_names_RVest.Rmd)
-   **Rob**
    -   Webscraping - [Tutorial](RU_scholar_names_RSelenium.Rmd)

`r colorize("BREAK", "red")`\
*Time: 12:30pm - 1:45pm*

`r colorize("genderize", "red")`\
*Time: 1:45pm - 2:45pm*

-   **Jochem**
    -   names curating - [tutorial](20names.html)
    -   names to gender - [tutorial](30gender.html)

`r colorize("Webscraping - publications", "red")`\
*Time: 2:45pm - 3:45pm*

-   **Jochem**
    -   Scraping publications - [tutorial](40publications.html)

`r colorize("Errors, bugs and crashes", "red")`\
*Time: 3:45pm - 4:30pm*

-   **All**
    -   Own work
    -   Questions
    -   Evaluation
