---
title: "Scraping RU Sociology department"
author: 
  - '**AUTHORS:**'
  - 'Rob Franken'
  - 'Daniel Cowen'  
  - 'Jochem Tolsma'
#bibliography: references.bib
link-citations: true
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
output: 
  html_document:
    css: tweaks.css
    toc:  true
    toc_float: true
    number_sections: true
    toc_depth: 2
    code_folding: show
    code_download: yes
---


```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(knitr)
library(tidyverse) 
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test3"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }
```

```{r klippy, echo=FALSE, include=TRUE}
#install.packages("remotes")
#remotes::install_github("rlesur/klippy")
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

# Preparation

## clean up
```{r, cleanup, results='hide'}
rm(list=ls())
gc()
```

<br>

## general custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R
- `fsave`: save data with time stamp in correct directory
- `fload`: load R-objects under new names


```{r, customfunc}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file, location = "./local/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}

fload  <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}

fshowdf <- function(x, digits = 2, ...) {
    knitr::kable(x, digits = digits, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}
```

<br>

## necessary packages

```{r, packages, results='hide', message=FALSE, warning=FALSE}
packages = c("RSelenium", "rvest", "tidyverse", "netstat", "pingr", "stringr")
fpackage.check(packages)
rm(packages)
```

<br>

## set up RSelenium

```{r, selenium, eval = FALSE}
#find a free port
port <- netstat::free_port(random = TRUE)

#ping port for confirmation
pingr::ping_port("localhost", port)

#set up selenium server and browser
rD <- rsDriver(browser="firefox", port=port, chromever = NULL, verbose = FALSE)
remDr <- rD[["client"]]
``` 

<br>

An issue with `RSelenium::rsDriver` is that it is configured to launch the Selenium server with the latest versions of Chrome *and* Firefox drivers, *even if you have specified the other browser to run*. To solve this, pass in `NULL` in the Chrome driver version `chromever`!

If at any time you wish to terminate the process, use the code:

```{r, taskkill, eval = FALSE}
pid <- rD$server$process$get_pid()#get process id
system(paste0("Taskkill /F /T" ," /PID ", pid))
```
<br>

---

# Scraping scholar names

```{r, names, eval = FALSE}
#navigate to RU staff page
remDr$navigate("https://www.ru.nl/en/search/scope/staff/")

#we want to get scholars from the sociology department; so we use the website's collapsible lists
#expand content under Radboud University
remDr$findElement(using = "class name", value = "facet-list__expand-sub")$clickElement()

#expand content under Faculty of social science (dep id 961)
#using its xpath
remDr$findElement(using = "xpath", value = "//button[@class='facet-list__expand-sub' and @aria-controls='toggle-staff-department-961']")$clickElement()

#click on sociology
remDr$findElement(using = "xpath", value = "//input[@type='checkbox' and @class='facets-checkbox' and @id='staff-department-992']")$clickElement()

#among all staff members, select researchers
#that is, check the 4th checkbox
remDr$findElements(using = "class name", value = "facets-checkbox")[[4]]$clickElement()

#naturally this can also be achieved by using the url https://www.ru.nl/en/search/scope/staff/staff-department/X/staff-staff/researchers, where X should be replaced department-id, which can be stored in a list beforehand

##################################################################################################

#now we have all sociology researchers;
#but this is spread across multiple pages;
#we store the elements of class "pager__item" in a list. We use this to get the number of pages (note that there are also "next" and "last" page buttons; so the number of elements of this class minus 2 equals the number of pages)

npages <- length(remDr$findElements("class name", "pager__item") ) - 2

#get all elements with class name "card__title"; these refer to names of researchers. Retrieve the text:
names_page <- unlist(lapply(remDr$findElements("class name", "card__title"),
                     function(x) x$getElementText()))

#now repeat this for the other pages in 1:npages-1
page1 <- remDr$getCurrentUrl()

for ( i in 1:(npages-1)) { 
  #increment page
  pageswitch <- paste0(page1,"?page=", i)
  #navigate to the page
  remDr$navigate(pageswitch)
  #retrieve names
  new_names <- unlist(lapply(remDr$findElements("class name", "card__title"),
                     function(x) x$getElementText()))
  #and add to the names_page object:
  names_page <- c(names_page, new_names)
  #add sleeper:
  Sys.sleep(2)
}

fshowdf(names_page)
``` 




