---  
title: "Logic of RSiena"
bibliography: references.bib
link-citations: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}

library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"), cache.lazy = FALSE)
options(width = 100) 
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```



```{r klippy, echo=FALSE, include=TRUE, message=FALSE}
# install.packages("remotes")
#remotes::install_github("rlesur/klippy")
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


----

# getting started

Start with clean workspace 

```{r}
rm(list=ls())
```

----

## Goal

Our goal is to understand the very basics of RSiena.  

For a thorough introduction in RSiena:  

- start with the [manual](https://www.stats.ox.ac.uk/~snijders/siena/RSiena_Manual.pdf)  
    - Chapters 2, 5, 12 (chapter 12 is the one I come back to most often)
- Follow the Rscripts [here](https://www.stats.ox.ac.uk/~snijders/siena/) or for .rmd style [here](https://jochemtolsma.github.io/RSiena-scripts/)  
- For an explanation how to interpret effects see [here](https://www.jochemtolsma.nl/courses/complete-networks/socio6/)
- Sign up for the mailing lists  
 

See this workshop more as a teaser. I hope that after the workshop you become enthusiastic about the logic of RSiena and are able to understand the manual. 


## Custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R ([source](https://vbaliga.github.io/verify-that-r-packages-are-installed-and-loaded/)).  
- `fsave`: Save to processed data in repository  
- `fload`: To load the files back after an `fsave`  
- `fshowdf`: To print objects (tibbles / data.frame) nicely on screen in .rmd  

```{r customfunctions, results='hide'}
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}

fsave <- function(x, file=NULL, location="./data/processed/") {
  ifelse(!dir.exists("data"), dir.create("data"), FALSE)
  ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
  if (is.null(file)) file= deparse(substitute(x))
  datename <- substr(gsub("[:-]", "", Sys.time()), 1,8)  
  totalname <- paste(location, datename, file, ".rda", sep="")
  save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
  load(filename)
  get(ls()[ls() != "filename"])
}

fshowdf <-  function(x, ...) {
  knitr::kable(x, digits=2, "html", ...) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::scroll_box(width="100%", height= "300px")
} 

```
## packages

- `RSiena`: what do you think?
- `RsienaTwoStep`: this packages assesses the ministep assumption of RSiena but is useful for this tutorial  
- `devtools`: to load from github  
- `igraph`: plotting tools. For a tutorial on plotting networks, see [our chapter in SNASS](https://snass.netlify.app/network-visualization.html) 


```{r, results='hide'}
packages = c("RSiena", "devtools", "igraph")
fpackage.check(packages)
#devtools::install_github("JochemTolsma/RsienaTwoStep", build_vignettes=TRUE)
packages = c("RsienaTwoStep")
fpackage.check(packages)
```

---  

# The logic of SAOMs

See [@ripley2022manual] paragraph 2.1

# RSiena as ABM 

RSiena models *the evolution* of network structures and/or the behavior of the social agents. 
It takes the current situation $T_0$ as starting point. It estimates the 'rules' for the agents how to change their ties and/or behavior. If the model is specified correctly, these rules have led the situation at $T_0$ to evolve into the situation observed at $T_1$. 

I would say these 'rules' are our ***micro theory of action***.  
Please note that our behavior may depend on the situation we are in. Similarly, the 'rules' we discover with RSiena are thus conditional on the situation at $T_0$. 

If we know the 'rules' of the social agents, we can also simulate future networks. And I think this aspect will help us to understand what the 'rules' of the social agents are and to understand what is estimated by `RSiena`. 

## RSiena's ministep

Before we can start to simulate or understand RSiena we need to know what is meant by the ministep assumption. 

Let us quote the manual 

> “The Stochastic Actor-Oriented Model can be regarded as an agent-based (‘actorbased’) simulation model of the network evolution; where all network changes are decomposed into very small steps, so-called ministeps, in which one actor creates or terminates one outgoing tie.” 
>
>“…it does not necessarily reflect a commitment to or belief in any particular theory of action elaborated in the scientific disciplines.”
>
>“It is assumed that actors change their scores on the dependent variable (tie or behavior) such that they improve their total satisfaction […] with her/his local network neighborhood configuration.”
>
>“Actors only evaluate all possible results in the local network neighborhood configurations that result from one ministep.”  

Thus, what does the SAOM of RSiena not do??:  

- No re-activity^[`r colorize("The no-reactivity assumption has been relaxed in the latest version of RSiena", "red")`]: The act of re-affirming, making or breaking an outgoing tie does not trigger a response by the involved alter  
- No simultaneity: Changes occur one by one  
- Hence also no cooperation, coordination or negotiation  
- No maximization of total utility:  
    - No altruistic behavior: Individual utility is maximized, not total utility  
- No strategic behavior:  
    - Very finite time horizon. Agent does not predict how his/her future local network neighborhood may change after:  
        - Making another ministep him/herself  
        - A ministep of other agents
    - Hence also no investments

This does not mean that RSiena cannot estimate (or better: 'fit') the evolution of networks/behavior that are the consequences of these more complex 'rules' or micro theories but it assumes actors only make ministeps.  



## Simulation Logic

1. Sample ego  
2. Construct possible alternative future networks based on all possible ministeps of ego    
3. Calculate how sampled ego evaluates these possible networks  
4. Let the ego pick a network, that is, let agent decide on a tie-change  
5. GOTO 1 (STOPPING RULE: until you think we have made enough ministeps)

### Sample an ego  

Let us first start with a network. We will use the build in network of `RsienaTwoStep`. 

#### net1

Adjacency matrix

```{r, echo=FALSE, eval=TRUE}
net1 <- net1
```


```{r }
net1
```

#### Plot net1

```{r, message=FALSE, warning=FALSE, fig.width=7, fig.height=5, fig.cap="net1"}
net1g <- graph_from_adjacency_matrix(net1, mode="directed")
coords <- layout_(net1g, nicely()) #let us keep the layout
par(mar=c(.1, .1, .1, .1))
{plot.igraph(net1g, layout=coords)
graphics::box()
}
```


So only one actor is allowed to make one ministep. But who? This is determined by the rate function and it may depend on ego-characteristics of our social agents (e.g. male/female) and/or on structural-characteristics of our social agents (e.g. indegree, outdegree). And all this can be estimated within RSiena. For this two hour course in RSiena it would go way too far to explain in detail the rate function. It is also not that important. More often than note, we simply assume that all actors have an equal chance of being selected to make a ministep.  

For more information on the rate function see [here](https://www.jochemtolsma.nl/courses/complete-networks/socio6/). 

#### select agent

Okay, we can thus randomly select an agent. 

```{r}
set.seed(24553253)
ego <- ts_select(net=net1, steps=1) #in rsienatwostep two actors may make a change together but here not
ego
```


### Possible networks after ministep  

Let us suppose we want to know what the possible networks are after all possible ministeps of `ego` who is part of `net1`. That is, let us assume that it is ego's turn (ego#: `r ego`) to decide on tie-change. What are the possible networks? 

The function `ts_alternatives_ministep()` returns a list of all possible networks after all possible tie-changes available to an ego given network da network.

```{r, results='hide'}
options <- ts_alternatives_ministep(net=net1, ego=ego)
options
plots <- lapply(options, graph_from_adjacency_matrix, mode="directed")
par(mar=c(0,0,0,0)+.1)
par(mfrow=c(2,2))

fplot <- function(x) {
  plot.igraph(x, layout=coords, margin=0)
  graphics::box()
}

lapply(plots, fplot) 
```



### Network statistics 

Which option will ego choose? Naturally this will depend on which network characteristics (or statistics) ego finds relevant. Let us suppose that ego bases its decision solely on the number of ties it sends to others and the number of reciprocated ties it has with others. 

Let us count the number of ties ego sends to alters. 

```{r}
ts_degree(net=options[[1]], ego=ego)
#or for all options

lapply(options, ts_degree, ego=ego)
```

And let us count the number of reciprocated ties

```{r}
lapply(options, ts_recip, ego=ego)
```


In the package `RsienaTwoStep` there are functions for the following network statistics $s$: 

  - degree: `ts_degree()` 
  - reciprocity: `ts_recip()`  
  - outdegree activity: `ts_outAct()` 
  - indegree activity: `ts_inAct()` 
  - outdegree popularity: `ts_outPop()` 
  - indegree popularity: `ts_inPop()`  
  - transitivity: `ts_transTrip()` 
  - mediated transitivity: `ts_transMedTrip()`  
  - transitive reciprocated triplets: `ts_transRecTrip()` 
  - number of three-cycles: `ts_cycle3()`  

`r colorize("Update: there are quite a lot more. But not very relevant for us at this time.", "red")`  

See @ripley2022manual (Chapter 12) for the mathematical formulation. Naturally, you are free to define your own network statistics. 

### Evaluation function 

But what evaluation value does ego attach to these network statistics and consequently to the network (in its vicinity) as a whole? Well these are the parameters, $\beta$, you will normally estimate with `RSiena::siena07()`. 
Let us suppose the importance for:  

- the statistic 'degree' is -1  
- for the statistic 'reciprocity' is 1.5. 

So you could calculate the evaluation of the network saved in `options[[4]]` by hand: 

$$f^{eval}_{n}(\mathbf{s_{i}},\mathbf{\beta}) = \mathbf{s_i}^\mathsf{T}\mathbf{\beta}, $$  
with $f^{eval}_{n}$ the evaluation function for agent $n$. $\mathbf{s_{i}}$ are the network statistics for network $i$ and $\mathbf{\beta}$ the corresponding parameters (or importance).

```{r}
option <- 4
ts_degree(options[[option]], ego=ego)*-1 + ts_recip(options[[option]], ego=ego)*1.5
```

Or you could use the `ts_eval()`. 

```{r}
eval <- ts_eval(net=options[[option]], ego=ego, statistics=list(ts_degree, ts_recip), parameters=c(-1,1.5))
eval
```
Now, let us calculate the evaluation of all possible networks: 
```{r t1, results='hold'}
eval <- sapply(options, FUN=ts_eval, ego=ego, statistics=list(ts_degree, ts_recip), parameters=c(-1,1.5))
eval
print("network with maximum evaluation score:")
which.max(eval)
```

### Choice function 

So which option will ego choose? Naturally this will be a stochastic process. But we see that option `r which.max(eval)` has the highest evaluation. 
We use McFadden's choice function (for more information see [wiki](https://en.wikipedia.org/wiki/Discrete_choice)), that is let $P_{ni}$ be the probability that ego $n$ chooses network/alternative $i$. The choice function is then given by:  


$$P_{n,i} = \frac{exp(\mathbf{s_i}^\mathsf{T}\mathbf{\beta} )}{\Sigma_{j=1}^J exp(\mathbf{s_j}^\mathsf{T}\mathbf{\beta} )},$$  

with $s_i$ a vector of the value of each network statistics for network $i$ and $\beta$ is the vector of parameter values. Hence, $\mathbf{s_i}^\mathsf{T}\mathbf{\beta}$ is the value of the evaluation for network $i$.

Let us force ego to make a decision. 

```{r, results='hold'}
choice <- sample(1:length(eval), size=1, prob=exp(eval)/sum(exp(eval)))
print("choice:") 
choice
#print("network:")
#options[[choice]]
```


If we repeat this process, that is...: 

1. sample agent  
2. construct possible alternative networks  
3. calculate how sampled agent evaluates the possible networks  
4. Let the agent pick a network, that is, let agent decide on a tie-change  
5. GO BACK TO 1 (STOPPING RULE: until you think we have made enough ministeps)

...we have an agent based model.   

### Stopping rule  

But how many ministeps do we allow? Well, normally this is estimated by `siena07` by the `rate` parameter.^[noterate: Naturally, it is a bit more complicated than that. In RSiena we have a choice between unconditional and conditional estimation. My description of the stopping rule refers to the unconditional estimation.] If we do not make this rate parameter conditional on actor covariates or on network characteristics, the rate parameter can be interpreted as the average number of ministeps each actor in the network is allowed to make before time is up. Let us suppose the `rate` parameter is 2 `r rate<-2`. Thus in total the number of possible ministeps will be `nrow(net1)*rate`: `r nrow(net1)*rate`. For a more detailed - **and more correct** -  interpretation of the rate parameter in `siena07` see: [www.jochemtolsma.nl](www.jochemtolsma.nl/courses/complete-networks/socio6).

## Simulation example  

Let us now simulate how the network **could** evolve given:^[It is also possible to simulate networks within the `RSiena` package itself. But let us stick with functions from `RsienaTwoStep` for now.]  

- starting point is `net1` 
- rate is set to 2  
- we as scientists think only network statistics degree and reciprocity are important  
- `RSiena::siena07` has determined the parameters for these statistics are -1 and 1.5 respectively  
- We adhere to the ministep assumption and hence set `p2step` to `c(1,0,0)`

```{r test, echo=FALSE, eval=TRUE}
data(net1)
net1
```

```{r}
ts_sims(nsims=1, net=net1, rate=2, statistics=list(ts_degree, ts_recip), parameters = c(-1,1.5), p2step=c(1,0,0), chain = FALSE)
```

# Estimation logic  

So now you have some grasp of the logic of RSiena. But how does the estimation work?  

Well, that is quite complicated. But it goes something like this:^['it' refers to estimation via the 'Method of Moments'] 

1. Define model: researcher includes effects  
2. initial parameter values of effects (commonly '0')  
3. simulate an outcome network based on these parameter values  
4. compare the network statistics of the simulated outcome network with the observed outcome network (i.e. the *target values*)  
    - based on the included effects. Thus the simulated network may contain 10 ties, but the observed network may contain 20 ties. Apparently, with the current parameter values we underestimate the density of the outcome network.  
5. tweak/update parameter values in some smart way  
6. GOTO 3 (BREAK RULE: until parameter values cannot be improved anymore / or reached good fit)  
7. simulate a bunch of outcome networks with the obtained parameter values and compare the *expected values* of statistics of the outcome networks with the *target values*.  
    - we can assess the fit  
    - estimate SE of the parameters




----

# References

