---  
title: "publication networks"
bibliography: references.bib
link-citations: yes
---
  
```{r, globalsettings, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"), cache.lazy = FALSE)
options(width = 100) 
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE, message=FALSE}
# install.packages("remotes")
#remotes::install_github("rlesur/klippy")
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


----

# getting started

Start with clean workspace 

```{r}
rm(list=ls())
```

----

# Goal

Our goal is to make a network based on the publications we scraped for our scholars. 


# Custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R ([source](https://vbaliga.github.io/verify-that-r-packages-are-installed-and-loaded/)).  
- `fsave`: Save to processed data in repository  
- `fload`: To load the files back after an `fsave`  
- `fshowdf`: To print objects (tibbles / data.frame) nicely on screen in .rmd  


```{r customfunctions, results='hide'}
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}

fsave <- function(x, file=NULL, location="./data/processed/") {
  ifelse(!dir.exists("data"), dir.create("data"), FALSE)
  ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
  if (is.null(file)) file= deparse(substitute(x))
  datename <- substr(gsub("[:-]", "", Sys.time()), 1,8)  
  totalname <- paste(location, datename, file, ".rda", sep="")
  save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
  load(filename)
  get(ls()[ls() != "filename"])
}

fshowdf <-  function(x, ...) {
  knitr::kable(x, digits=2, "html", ...) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::scroll_box(width="100%", height= "300px")
} 

```


# packages

- `stringdist`: string stuff.
- `stringi`: string stuff 

```{r, results='hide'}
packages = c("stringdist", "stringi", "tidyverse")

fpackage.check(packages)
```

# load data


Hopefully you managed to construct the datasets yourself. If not. Here are the downloads: 

Please use this one:  

`r xfun::embed_file("./data/processed/20230621df_complete.rda")`. 

`r xfun::embed_file("./data/processed/20230621list_publications_jt.rda")`. 

Save file in correct directory: './data/processed'. 




```{r}
df <- fload("./data/processed/20230621df_complete.rda")
publications <- fload("./data/processed/20230621list_publications_jt.rda")
```

## unique publications in df

```{r}
publications <- publications %>% 
  bind_rows() %>%
  distinct(title, .keep_all = TRUE) 
```

---  

# networks based on publications  

We have to make a couple of decisions:  

- what are the boundaries of our network?: 
  - combination of university and discipline  
- undirected or directed network?:  
  - directed: first author is sending a tie to other authors  
- weighted or unweighted network?:  
  - unweighted (no distinction between number ties within a specific time window, thus 0/1)  
- time window?: 
  - one wave is 3 years   
  - 2 waves  
    - wave 1: 2018-2020  
    - wave 2: 2021-2023  

## An example: RU-sociology

```{r}
#select scholars
df %>% 
  filter(affil1=="RU" | affil2=="RU") %>%
  filter(discipline=="sociology") -> df_sel
fshowdf(df_sel)
```

### network based on publications

We will use an adjacency matrix to store our network ties: the first author is sending a tie to other authors. 

We make the assumption that the composition of the network is stable!  

```{r}
networkw1 <- matrix(0, nrow=nrow(df_sel), ncol=nrow(df_sel))
networkw2 <- matrix(0, nrow=nrow(df_sel), ncol=nrow(df_sel))
```


### select publications

```{r}
#wave1
publications %>% 
  filter(gs_id %in% df_sel$gs_id) %>% 
  filter (year %in% c(2018,2019,2020))-> pub_w1

#wave2
publications %>% 
  filter(gs_id %in% df_sel$gs_id) %>% 
  filter (year %in% c(2021,2022,2023))-> pub_w2
```

### make a list for each publication 

cleaned a bit
```{r}
#wave1
pub_w1 <- str_split(pub_w1$author, ",") %>%
  #lowercase
  lapply(tolower) %>% 
  # Removing diacritics
  lapply(stri_trans_general, id = "latin-ascii") %>%
  # only last name
  lapply(word, start=-1, sep=" ") %>%
  # only last last name
  lapply(word, start=-1, sep="-")
 
#let us remove all publications with only one author
remove <- which(sapply(pub_w1, FUN = function(x) length(x)==1 )==TRUE)
pub_w1 <- pub_w1[-remove]

#wave2
pub_w2 <- str_split(pub_w2$author, ",") %>%
  #lowercase
  lapply(tolower) %>% 
  # Removing diacritics
  lapply(stri_trans_general, id = "latin-ascii") %>%
  # only last name
  lapply(word, start=-1, sep=" ") %>%
  # only last last name
  lapply(word, start=-1, sep="-")
 
#let us remove all publications with only one author
remove <- which(sapply(pub_w2, FUN = function(x) length(x)==1 )==TRUE)
pub_w2 <- pub_w2[-remove]
```

### find the author positions  

And if we know the author positions, we can fill in our matrix  
```{r}
pubs <- pub_w1
for (ego in 1: nrow(df_sel)) {
  #which ego? 
  lastname_ego <- df_sel$lastname[ego]
  #for all publications
  for (pub in 1:length(pubs)) {
    #only continue if ego is author of pub
    if (lastname_ego %in% pubs[[pub]]) {
      aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
      #only continue if ego is first author of pub
      if (aut_pot==1) {
        #check all alters/co-authors
        for (alter in 1: nrow(df_sel)) {
          #which alter
          lastname_alter <- df_sel$lastname[alter]
          if (lastname_alter %in% pubs[[pub]]) {
            networkw1[ego,alter] <- networkw1[ego,alter] + 1
          }
        }
      }
    }
 }
}

pubs <- pub_w2

for (ego in 1: nrow(df_sel)) {
  #which ego? 
  lastname_ego <- df_sel$lastname[ego]
  #for all publications
  for (pub in 1:length(pubs)) {
    #only continue if ego is author of pub
    if (lastname_ego %in% pubs[[pub]]) {
      aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
      #only continue if ego is first author of pub
      if (aut_pot==1) {
        #check all alters/co-authors
        for (alter in 1: nrow(df_sel)) {
          #which alter
          lastname_alter <- df_sel$lastname[alter]
          if (lastname_alter %in% pubs[[pub]]) {
            networkw1[ego,alter] <- networkw1[ego,alter] + 1
          }
        }
      }
    }
 }
}
```


---  

# making life easy

Wouldn't it be great if we could not write a function that would all this in one go? 

Okay

```{r}
f_pubnets <- function(df_scholars=df, list_publications=publications, discip="sociology", affiliation="RU", waves=list(wave1=c(2018,2019,2020), wave2=c(2021,2022,2023))) {
  
  publications <- list_publications %>% 
  bind_rows() %>%
  distinct(title, .keep_all = TRUE) 
  
  df_scholars %>% 
  filter(affil1==affiliation | affil2==affiliation) %>%
  filter(discipline==discip) -> df_sel
  
  networklist <- list()
  for (wave in 1:length(waves)) {
    networklist[[wave]] <- matrix(0, nrow=nrow(df_sel), ncol=nrow(df_sel))
  }
  
  publicationlist <- list()
  for (wave in 1:length(waves)) {
    publicationlist[[wave]] <- publications %>%
    filter(gs_id %in% df_sel$gs_id) %>% 
    filter (year %in% waves[[wave]]) %>%
    select(author) %>% 
    lapply(str_split, pattern=",") 
  }  
    
  publicationlist2 <- list()
  for (wave in 1:length(waves)) {
    publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
    #lowercase
    lapply(tolower) %>% 
    # Removing diacritics
    lapply(stri_trans_general, id = "latin-ascii") %>%
    # only last name
    lapply(word, start=-1, sep=" ") %>%
    # only last last name
    lapply(word, start=-1, sep="-")
  }
  
  for (wave in 1:length(waves)) {
    #let us remove all publications with only one author
    remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x)==1 )==TRUE)
    publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
  }

  for (wave in 1:length(waves)) {
    pubs <- publicationlist2[[wave]]
    for (ego in 1: nrow(df_sel)) {
      #which ego? 
      lastname_ego <- df_sel$lastname[ego]
      #for all publications
      for (pub in 1:length(pubs)) {
        #only continue if ego is author of pub
        if (lastname_ego %in% pubs[[pub]]) {
          aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
          #only continue if ego is first author of pub
          if (aut_pot==1) {
            #check all alters/co-authors
            for (alter in 1: nrow(df_sel)) {
              #which alter
              lastname_alter <- df_sel$lastname[alter]
              if (lastname_alter %in% pubs[[pub]]) {
                networklist[[wave]][ego,alter] <- networklist[[wave]][ego,alter] + 1
              }
            }
          }
        }
      }
    }
  }
  return(list(df=df_sel, network=networklist))
}
```

## some testing

```{r}
outputUU_sociology <- f_pubnets(affiliation="UU")
str(outputUU_sociology)
head(outputUU_sociology[[1]])
#head(outputUU_sociology[[2]])
```

It seems to work. 

