---
title: "Sunbelt 2023 Workshop: from webscraping to SNA"
author: 
  - '**AUTHORS:**'
  - 'Rob Franken'  
  - 'Anne Maaike Mulders' 
  - '[Bas Hofstra](https://www.bashofstra.com/)'
  - '[Jochem Tolsma](https://www.jochemtolsma.nl) - Radboud University / University of Groningen, the Netherlands'
date: "Last compiled on `r format(Sys.Date(), format='%d %B %Y')`"
bibliography: references.bib
link-citations: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r, globalsettings, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"), cache.lazy = FALSE)
options(width = 100) 
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE, message=FALSE}
# install.packages("remotes")
#remotes::install_github("rlesur/klippy")
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

------------------------------------------------------------------------

# Introduction

```{=html}
<script>
function myFunction() {

            var btn = document.getElementById("myButton");
            //to make it fancier
            if (btn.value == "Click to hide code") {
                btn.value = "Read more";
                btn.innerHTML = "Read more";
            }
            else {
                btn.value = "Click to hide code";
                btn.innerHTML = "Read less";
            }
            //this is what you're looking for
            var x = document.getElementById("myDIV");
            if (x.style.display === "none") {
                x.style.display = "block";
            } else {
                x.style.display = "none";
            }
        }
          
</script>
```
Welcome to the website of the Sunbelt 2023 Workshop: **From Webscraping
to Social Network Analyses: Structure and Evolution of Scientific
Co-publishing Networks**

<br>

In this workshop you will first collect and then analyze academic
departmental co-publishing networks. In the workhop's Part 1 you learn
to webscrape scientific metadata of scientific departmental websites
(via R packages like rvest and RSelenium), assign name-based gender and
ethnicity signals, retrieve scholars' publications, and construct
longitudinal co-publishing networks. In Part 2, you learn to analyze the
structure and evolution of these networks by means of RSiena. You will
answer questions such as: are co-publishing networks segregated by
scientific success? And is success-based segregation in co-publishing
networks the result of departmental characteristics, structural network
effects, influence processes and selection processes? For each step, we
provide clear (proof-of-principle) coding examples and output data,
ensuring you will not get stuck along the way. Depending on skills and
progress, you might analyze your own chosen departments. You will keep
track of your work via a labjournal on GitHub.

<!--

<button class="button1" onclick="myFunction()" id="myButton" value="Click To Open Instructions">

Read more

</button>

::: {#myDIV style="display:none;"}
<br>

more text here

<br>
:::

--->

------------------------------------------------------------------------

# Getting started

## LabJournal

During the course, students will journal their work and assignments in
their custom lab journal. A template lab journal can be found on
[GitHub](https://github.com/robfranken/LabJournal). Here, you find how
to get started.

## Discord

Please join the [discord channel](https://discord.gg/ACcYU8nu). Use the
channel to chat, ask questions and share (small) files.

------------------------------------------------------------------------

# Program

The program of this workshop will be as follows:

`r colorize("Introduction", "red")`\
*Time: 9:00am - 9:30am*  

- **all**  
    - round of introductions  
- **Jochem and Anne Maaike**
    -   Our RQs for today\
    -   Data requirements for RQs\
    -   PhD project AM

`r colorize("Lab Journal", "red")`\
*Time: 9:30am - 10:00am*

-   **Rob**
    -   [Lab Journal](https://github.com/robfranken/LabJournal)

`r colorize("Webscraping-scholars", "red")`\
*Time: 10:00am - 11:30am*

-   **Bas**
    -   Webscraping - [Lecture](./presentations/insna_science_data.pdf)
    -   Webscraping - [Tutorial](10scholars.html)

`r colorize("BREAK", "red")`\
*Time: 11:30am - 11:45am*

`r colorize("genderize", "red")`\
*Time: 11:45am - 12:30pm*

-   **Anne Maaike**
    -   names curating - [tutorial](20names.html)
    -   names to gender - [tutorial](30gender.html)

`r colorize("BREAK", "red")`\
*Time: 12:30pm - 1:45pm*

`r colorize("Webscraping - publications", "red")`\
*Time: 1:45pm - 2:00pm*

-   **Jochem and Bas**
    -   Scraping publications - [demonstration](40publications.html)

`r colorize("RSiena", "red")`\

*Time: 2:00pm - 4:00pm*

-   **Jochem and Rob**
    -   Logic of RSiena - [lecture](./presentations/insna_logic_rsiena.pdf)  
    -   Rsiena data preparation - [tutorial](50pubnets.html)  
    -   RSiena analyses  - [tutorial](60RSiena.html)  
    

`r colorize("The End", "red")`\
*Time: 4:00pm - 4:30pm*

-   **All**
    -   Questions  
    -   Evaluation  

---  

